\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Vishnu V Chandrabalan}
\begin{document} 
		
\section{Chapter 1 - More discussion of key RCTs}
		
The following paragraph has been included as Section 1.x in Chapter 1.
		
\section{Chapter 1 - Short discussion of aetiology needed}
		
\section{Chapter 2 - Review patient numbers on page 63}

Patient numbers have been updated to account for the missing information and this section now reads as below accounting for the 100 patients included in this study.

\textit{Pathological examination of the resected specimen showed pancreatic ductal adenocarcinoma (n=37), ampullary adenocarcinoma (n=18), cholangiocarcinoma (n=17), duodenal adenocarcinoma (n=6), intra-ductal papillary mucinous neoplasia (n=4), neuroendocrine tumours (n=7), other neoplasia (n=6) or chronic pancreatitis (n=5).}

\section{Chapter 2 - Aim and end points must match}

The description of the aims of this chapter has been updated to match the end-points described later in methods and results sections and now reads as below.

\textit{The aim of the present study was to evaluate the role of various measures of patient physiological fitness including cardiopulmonary exercise testing in predicting postoperative length of stay, major postoperative adverse events including operative mortality and fitness to undergo adjuvant therapy when indicated after pancreaticoduodenectomy.}

\section{Page 69 - Error in AT numbers}

The results presented in this table have been reanalysed using the primary database and have been found to be accurate. In the group of patients where adjuvant chemotherapy was indicated, anaerobic threshold was low in a greater proportion. However, low anaerobic threshold was associated with non-receipt of adjuvant therapy in spite of being indicated as shown in this table.

\section{Page 72 - Remove unjustified/unsupported conclusion, Page 73 - Remove GPS conclusion}

I am not entirely sure how to address this. Should I completely remove this sentence? 

The following sentence has been removed from the discussion in Chapter 2. 

\textit{Therefore, it would appear that objective measurement of patient physiological fitness using cardiopulmonary exercise testing is superior to conventional measures of patient fitness including the POSSUM Physiology Score or the modified Glasgow Prognostic Score and may have a role in predicting short-term outcome which in turn affects the overall management of these patients including receipt of adjuvant therapy.}

\section{Chapter 3, page 96 - Remove conclusions not based on thesis}

The following sentence has been removed from the discussion in Chapter 3.

\textit{Taken together, these results appear to suggest that preoperative biliary drainage and resolution of jaundice are unlikely to improve aerobic capacity and cardiopulmonary physiology in these patients.}


\section{Chapter 5,6 - Reduce to more understandable key issues. Several irrelevant tables should be removed.}

\section{All acronyms/abbreviations on list}

A list of abbreviations has been included. 

\section{All tables should have units (e.g. mg/L, etc) and legends.}

All tables have been updated to include units. A brief description has been added as a legend to all figures and tables. 

\section{All values and ranges need defining}
			
		
\end{document}